
\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{unicode-math}
\usepackage{fancyhdr}
\usepackage[margin=1.0in, left=0.5in, right=0.5in, top=1.25in, headsep=10mm, headheight=15mm]{geometry}
\usepackage{graphicx}

\setmainfont{Latin Modern Roman}
\setmathfont{Latin Modern Math}

\pagestyle{fancy}
\rhead{Mech 222 Mathematics Worksheets\\ Lecture 16, 17}

\input{preamble_common.tex}

\begin{document}
\section*{Surface Parameterization}
There are two ways of writing a surface in parametric form:
\begin{enumerate}
\item Writing it in vector form, ie $r(s, t) = x(s, t) \hat{i} + y(s, t) \hat{j} + z(s, t) \hat{k}$. This is the parametric form.
\item Solve the equation for one of the variables, $z = f(x, y)$.
  This is a special case of the parametric form $r(s, t) = s \hat{i} + t \hat{j} + f(s, t) \hat{k}$, it's called the graph form of the surface.
\end{enumerate}
Parameterizing a surface is similar to parameterizing a curve; only with an extra dimension to consider.
Similar tricks apply; remembering cylindrical and spherical coordinates can be very helpful!
For planes, it's best to find two linearly independent vectors which are part of the plane.

\section*{Surface Area}
In parametric form $r(s, t)$, the surface area can be computed as $\iint_D |r_u \times r_v| \dd{A}$.
In graph form, the surface area can be computed as $\iint_D \sqrt{1 + \left(\pderiv{z}{x}\right)^2 + \left(\pderiv{z}{y}\right)^2} \dd{A}$

\section*{Problems}
\begin{enumerate}
\item Parameterize the plane defined by $4 x + 3 y - 2 z = 3$.
  Compute the surface area over the square $0 \leq x \leq 1$, $0 \leq y \leq 1$.
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
\item Write in parametric form the surface of a cone of height $1$ and radius equal to the height.
  Compute the surface area.
  \newpage
\item Parameterize the surface defined by $4 x^2 + \frac{1}{3} y^2 - z = 0$.
  Write the integral (don't compute it) for the surface area for $0 \leq z \leq 1$.
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
  \newline
\item {\bf Challenge:} Parameterize the surface defined by $z^2 + x^2 + y^2 + x y = 1$.
  Write the integral (don't compute it) for the surface area.\\
  Hint: Try rotating the x-y coordinate system by $\pm 45$ degrees.
  A rotation matrix in 2D is given by
  $$
  \begin{pmatrix}
    \cos(\theta) & \sin(\theta)\\
    -\sin(\theta) & \cos(\theta)
  \end{pmatrix}
  $$
  \newpage
\end{enumerate}
\end{document}
